%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title: SOP LaTeX Template
%% Author: Soonho Kong / soonhok@cs.cmu.edu
%% Created: 2012-11-12
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Requirement:
%%     You need to have the `Adobe Caslon Pro` font family.
%%     For more information, please visit:
%%     http://store1.adobe.com/cfusion/store/html/index.cfm?store=OLS-US&event=displayFontPackage&code=1712
%%
%% How to Compile:
%%     $ xelatex main.tex
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[letterpaper]{article}
\usepackage[letterpaper,margin=1.75in,noheadfoot]{geometry}
\usepackage{fontspec, color, enumerate, sectsty}
\usepackage[normalem]{ulem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                      YOUR INFORMATION
%
%      PLEASE EDIT THE FOLLOWING LINES ACCORDINGLY!!
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\soptitle}{Statement of Purpose}
\newcommand{\yourname}{Jeffrey Josanne Michael}
\newcommand{\youremail}{jeffreyj@cse.iitb.ac.in}

\usepackage[bookmarks, colorlinks, breaklinks,
pdftitle={\yourname - \soptitle},pdfauthor={\yourname}, unicode]{hyperref}
\hypersetup{linkcolor=magneta,citecolor=magenta,filecolor=magenta,urlcolor=[named]{WildStrawberry}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                      Title and Author Name
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{center}{\huge \scshape \soptitle}\end{center}
\begin{center}\vspace{0.2em} {\Large \yourname\\}
  {\youremail}\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                      SOP Body
% NOTE: Use \amper instead of \&
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{\par}
I will let my kids write about me. My first kid, I mean the \textbf{text summarization script} I wrote to shorten my detailed purpose statement with summary ratio of 94.01\%. The summarized text: \textit{ \textbf{‘He demonstrated his aptitude for research in internships and projects where he quickly learnt new tools and web technologies’}}. My second kid, the markov , [Made pronoun (‘I’ to ‘He’ and ‘my’ to ‘his’) changes to the corpus].\\ \par

During my undergraduation I had the honor of being selected to lead the Web development team of my college’s official systems. Furthermore, Dr. Diwan Baskaran’s guidance and the loads of data present in the system, I developed, tempted me to use Latent Dirichlet Allocation and K-means clustering to categorize questions and rank answers using grammar models (pre-trained). Later I worked under Dr. Sumathi to assist with her PhD research, which creates a Clause-level analysis by tuning the grammar model intrinsically. With relevant experience, I was hired as a research intern by Cognizant Technology Solutions R\&D department to design and conduct in-depth analysis of a Speech recognition engine in which I added distributions to the syntactic fields in acoustic model for regional English accents. This helped increase the efficiency before the implementation of Hidden Markov Model level itself. Now, I am at Indian Institute of Technology, Bombay (IIT-B) researching on content (video) categorization from a hierarchical model to Pachinko allocation model. This project has future prospects of extending my research to video summarization by building a convolutional neural network if required. I have actively worked on many web developments and software developments also (refer résumé). I have honed my leadership skills from the given opportunities. For instance, I have been appointed as a Campus ambassador for Indian Institute of Management Bangalore for managing our college’s team for proposing business ideas which are backed by strong technical and commercial potential. Currently, I am leading a team of 7 research interns in Indian Institute of Technology.  Having specialized in Natural Language Processing (NLP) and Machine Learning (ML) research  under 3 professors and one R\&D department, I crave to work on the exciting projects under the professors at University of Illinois at Urbana-Champaign (UIUC).\\ \par


My relationship with NLP is personal. When I reached IIT Bombay for my research, I was totally alien to Hindi, the language used by most of the people here. Being a NLP researcher, I planned to learn Hindi the way an artificial Neural Network (NN) would learn a new language. We train an infant NN by exposing it to a history of experiences (datasets), and this model of perceptrons starts understanding context, grammar and framing sentences. In the same way, every time I heard Hindi, I got used to the context by Bag of Words (BoW) (mentally, remembering meanings of some common words). I was even able to train myself to form the language model (mentally) with n-grams. As I got familiar and fast paced with BoW, I was able to train myself to skip words to learn contexts considering collocations and using Skip-gram technique. Now I'm leading a team of people who speak Hindi. Thanks to NLP. \\ \par


The reason for me applying to Masters in the United States is that research projects are funded regularly, allowing researchers to concentrate on research. Having gone through both rewarding and frustrating experiences during my relatively short (2.5 years) research career, I feel confident that I would like to pursue research in the field of NLP. During my research at UIUC, I will be thoroughly interested in taking a project from Lambda calculus (theories) to scripts (implementation), by which I mean I am eager to participate actively in the theoretical modelling of the research and also the implementation of the Professor's research.\\ \par


As I will be working in Video summarization in IIT-B before I start the Fall semester at UIUC, Dr. Julia Hockenmaier’s research on \textit{‘Describing images with sentences’} will help me gain extensive knowledge in the same domain and also I can contribute to her research with my experience at IIT-B. And also, while analysing a scientific hypothesis especially in NLP, I have always approached a problem in a syntactic manner (logically) and have never been prone to a semantic approach. But there are complexities in languages when the model is only syntactically designed. Dr. Julia’s research on \textit{‘Combinatory Categorial Grammar parsing’} has both syntactic and semantic approaches to the same. This will expose me to another research perspective. Dr. Julia’s  NLP group and Dr. Dan Roth’s Cognitive computation group have discussions every thursdays and I feel that the synergy of these two major concepts will dominate Intelligent systems in the future.\\ \par


I intend to pursue my Masters degree at your university as I feel my area of interests go in tune with NLP research at UIUC powered by five diverse research groups and make major contributions to the NLP \& ML research community. This would allow me to fulfill my long term goal of transforming the behaviour of any model from \textit{Perception-Action} to \textit{Perception-Cognition-Action}. I have plans to accomplish it and then to create such frameworks that can be used by any computing model which perceive and perform computations. Having my research goal as so, my career goal is to become a researcher at productive and intellectually rewarding research environment. I see myself mentoring new researchers and manage projects down the road. I feel excited applying to UIUC especially because of the five different NLP research groups. I would have the flexibility of discussing with the group members and finalize my research thesis from various options in NLP.\\ \par


Frequent discussions with graduate students at UIUC and  my own research about the university backed by my intuition have convinced me that research and exhaustive coursework being conducted at UIUC is ideally suited for my ambitions and I am certain that it will serve as an efficient catalyst for me to strive hard and achieve success.

\end{document}
